A Large-scale Study of Polymorphism in R

### 2019-03-28 paper/main.tex
more data, more discussion

\newcommand{\PACKAGES}{11,463\xspace}
\newcommand{\PROGRAMMERS}{?\xspace}
\newcommand{\PERCENTCRAN}{83\%\xspace}
\newcommand{\CRANTOTAL}{13,841\xspace}
\newcommand{\RLOC}{15,050,267\xspace}
\newcommand{\CLOC}{9,373,542\xspace}
\newcommand{\YEARS}{?\xspace}
\newcommand{\INDEXCOINCIDENCE}{3,426\xspace}
\newcommand{\TOTALINDEXY}{6,461\xspace}
\newcommand{\INDEXYPERC}{53\%\xspace}		% TODO Not an accurate count.

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:03:30 2019
\begin{table}[ht]
\label{tab:argcountsFGTS}
\centering
\begin{tabular}{lrr}
  \hline
 Argument Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 780K & 85.36 \\ 
  Monomorphic in Type & 832K & 91.03 \\ 
  Monomorphic in Class & 827K & 90.41 \\ 
  Monomorphic in Attribute Pattern & 825K & 90.27 \\ 
  Total Seen & 914K & --- \\ 
     \hline
\end{tabular}
\caption{Account of {\it argument} polymorphism in the FGTS.}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:10:46 2019
\begin{table}[ht]
\label{tab:funcountsFGTS}
\centering
\begin{tabular}{lrr}
  \hline
 Function Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 143K & 65.87 \\ 
  Monomorphic in Type & 167K & 76.95 \\ 
  Monomorphic in Class & 163K & 75.21 \\ 
  Monomorphic in Attribute Pattern & 161K & 74.37 \\ 
  Total Seen & 216K & --- \\ 
   \hline
\end{tabular}
\caption{Account of {\it function} polymorphism in the FGTS.}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 21:27:37 2019
\begin{table}[ht]
\label{tab:toppolyFDTS}
\centering
\begin{tabular}{lrr}
  \hline
 Polymorphic Type Signature & \# Occurrences & \% of Total Polymorphism \\ 
  \hline
  vector/double, vector/integer & 8228 & 10.03 \\ 
  scalar/double, scalar/integer & 7190 & 8.76 \\ 
  matrix, vector/double & 5362 & 6.54 \\ 
  NULL, scalar/character & 2308 & 2.81 \\ 
  list$<$any$>$, list$<$double$>$ & 2220 & 2.71 \\ 
  data.frame, matrix & 2115 & 2.58 \\ 
  NULL, scalar/double & 2064 & 2.52 \\ 
  NULL, vector/double & 2041 & 2.49 \\ 
  list$<$any$>$, list$<$list$>$ & 1420 & 1.73 \\ 
  NULL, vector/character & 1374 & 1.67 \\ 
     \hline
\end{tabular}
\caption{Top polymorphic argument signatures in the FDTS.}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:03:30 2019
\begin{table}[ht]
\label{tab:argcountsWHATTS}
\centering
\begin{tabular}{lrr}
  \hline
 Argument Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 781K & 85.46 \\
  Monomorphic in Type & 867K & 94.78 \\
%  Monomorphic in Class & 831K & 90.31 \\ 
%  Monomorphic in Attribute Pattern & 830K & 90.25 \\ 
  Total Seen & 914K & --- \\ 
     \hline
\end{tabular}
\caption{Account of {\it argument} polymorphism in the NEW TYPE SYSTEM.}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:10:46 2019
\begin{table}[ht]
\label{tab:funcountsWHATTS}
\centering
\begin{tabular}{lrr}
  \hline
 Function Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 143K & 66.02 \\ 
  Monomorphic in Type & 184K & 84.91 \\
%  Monomorphic in Class & 163K & 75.13 \\ 
%  Monomorphic in Attribute Pattern & 161K & 74.31 \\ 
  Total Seen & 216K & --- \\ 
   \hline
\end{tabular}
\caption{Account of {\it function} polymorphism in the NEW TYPE SYSTEM.}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Tue Mar 26 17:46:45 2019
\begin{table}[ht]
\label{tab:toppolyWHATTS}
\centering
\begin{tabular}{lrr}
  \hline
Polymorphic Type Signature & \# Occurrences & \% of Total Polymorphism \\ 
  \hline
  matrix, vector/real & 6320 & 13.25 \\ 
  list$<$any$>$, list$<$real$>$ & 2782 & 5.83 \\ 
  vector/character, vector/real & 2706 & 5.67 \\ 
  scalar/character, scalar/real & 2523 & 5.29 \\ 
  data.frame, matrix & 2163 & 4.53 \\ 
  list$<$any$>$, list$<$list$>$ & 1484 & 3.11 \\ 
  list$<$any$>$, list$<$character$>$ & 1211 & 2.54 \\ 
  data.frame, vector/real & 1027 & 2.15 \\ 
  scalar/logical, scalar/real & 911 & 1.91 \\ 
  scalar/character, vector/real & 839 & 1.76 \\ 
   \hline
\end{tabular}
\caption{Top 10 argument signatures in the NEW TYPE SYSTEM.}
\end{table}

@@

This figure reveals an unfortunate truth that date ``data type'' in R is not standardized.
That said, translating between the data types is straigthforward.

@@

>>>>

\newcommand{\PACKAGES}{11,463\xspace}
\newcommand{\PROGRAMMERS}{?\xspace}
\newcommand{\PERCENTCRAN}{83\%\xspace}
\newcommand{\CRANTOTAL}{13,841\xspace}
\newcommand{\RLOC}{15,050,267\xspace}
\newcommand{\CLOC}{9,373,542\xspace}
\newcommand{\YEARS}{?\xspace}
\newcommand{\INDEXCOINCIDENCE}{3,499\xspace}
\newcommand{\TOTALINDEXY}{6,561\xspace}
\newcommand{\INDEXYPERC}{53\%\xspace}		% TODO Not an accurate count.

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:03:30 2019
\begin{table}[ht]
\label{tab:argcountsFGTS}
\centering
\begin{tabular}{lrr}
  \hline
 Argument Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 781K & 85.37 \\ 
  Monomorphic in Type & 834K & 91.23 \\ 
  Monomorphic in Class & 827K & 90.42 \\ 
  Monomorphic in Attribute Pattern & 825K & 90.27 \\ 
  Total Seen & 914K & --- \\ 
     \hline
\end{tabular}
\caption{Account of {\it argument} polymorphism in the FGTS.}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:10:46 2019
\begin{table}[ht]
\label{tab:funcountsFGTS}
\centering
\begin{tabular}{lrr}
  \hline
 Function Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 143K & 65.88 \\ 
  Monomorphic in Type & 167K & 77.25 \\ 
  Monomorphic in Class & 163K & 75.24 \\ 
  Monomorphic in Attribute Pattern & 161K & 74.37 \\ 
  Total Seen & 216K & --- \\ 
   \hline
\end{tabular}
\caption{Account of {\it function} polymorphism in the FGTS.}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 21:27:37 2019
\begin{table}[ht]
\label{tab:toppolyFDTS}
\centering
\begin{tabular}{lrr}
  \hline
 Polymorphic Type Signature & \# Occurrences & \% of Total Polymorphism \\ 
  \hline
  vector/double, vector/integer & 7742 & 9.66 \\ 
  scalar/double, scalar/integer & 7201 & 8.99 \\ 
  matrix, vector/double & 5384 & 6.72 \\ 
  NULL, scalar/character & 2325 & 2.90 \\ 
  list$<$any$>$, list$<$double$>$ & 2231 & 2.78 \\ 
  data.frame, matrix & 2116 & 2.64 \\ 
  NULL, scalar/double & 2082 & 2.60 \\ 
  NULL, vector/double & 2069 & 2.58 \\ 
  list$<$any$>$, list$<$list$>$ & 1432 & 1.79 \\ 
  NULL, vector/character & 1380 & 1.72 \\
     \hline
\end{tabular}
\caption{Top polymorphic argument signatures in the FDTS.}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:03:30 2019
\begin{table}[ht]
\label{tab:argcountsWHATTS}
\centering
\begin{tabular}{lrr}
  \hline
 Argument Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 781K & 85.47 \\
  Monomorphic in Type & 867K & 95.01 \\
%  Monomorphic in Class & 831K & 90.31 \\ 
%  Monomorphic in Attribute Pattern & 830K & 90.25 \\ 
  Total Seen & 914K & --- \\ 
     \hline
\end{tabular}
\caption{Account of {\it argument} polymorphism in the NEW TYPE SYSTEM.}
\end{table}

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Fri Mar 22 18:10:46 2019
\begin{table}[ht]
\label{tab:funcountsWHATTS}
\centering
\begin{tabular}{lrr}
  \hline
 Function Polymorphism & Count & Percentage \\ 
  \hline
  Full Monomorphic & 143K & 66.04 \\ 
  Monomorphic in Type & 184K & 85.30 \\
%  Monomorphic in Class & 163K & 75.13 \\ 
%  Monomorphic in Attribute Pattern & 161K & 74.31 \\ 
  Total Seen & 216K & --- \\ 
   \hline
\end{tabular}
\caption{Account of {\it function} polymorphism in the NEW TYPE SYSTEM.}
\end{table}

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Tue Mar 26 17:46:45 2019
\begin{table}[ht]
\label{tab:toppolyWHATTS}
\centering
\begin{tabular}{lrr}
  \hline
Polymorphic Type Signature & \# Occurrences & \% of Total Polymorphism \\ 
  \hline
  matrix, vector/real & 6360 & 13.94 \\ 
  list$<$any$>$, list$<$real$>$ & 2799 & 6.13 \\ 
  vector/character, vector/real & 2774 & 6.08 \\ 
  scalar/character, scalar/real & 2552 & 5.59 \\ 
  data.frame, matrix & 2167 & 4.75 \\ 
  list$<$any$>$, list$<$list$>$ & 1496 & 3.28 \\ 
  list$<$any$>$, list$<$character$>$ & 1214 & 2.66 \\ 
  data.frame, vector/real & 1046 & 2.29 \\ 
  scalar/logical, scalar/real & 920 & 2.02 \\ 
  scalar/character, vector/real & 868 & 1.90 \\
   \hline
\end{tabular}
\caption{Top 10 argument signatures in the NEW TYPE SYSTEM.}
\end{table}

@@

This figure reveals an unfortunate truth that date ``data type'' in R is not standardized.
That said, translating between the data types is straightforward.

@@

Which packages are the most polymorphic?
Figure~\ref{tab:packagepolysize} has the answer.

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Thu Mar 28 18:15:26 2019
\begin{table}[ht]
\label{tab:packagepolysize}
\centering
\begin{tabular}{lrl}
  \hline
Package & Total Size & Description \\ 
  \hline
  magrittr & 4.16 & Tidyverse pipe suite \\ 
  types & 3.25 & Simple type annotations for R \\ 
  r6extended & 3.00 & Utilities extending barebones R6 package \\ 
  utils & 2.93 & Utilities for R developers \\ 
  zeallot & 2.85 & Metaprogramming package for assignment \\ 
  listviewer & 2.72 & R list visualizer \\ 
  proto & 2.60 & Prototype object-based programming \\ 
  apercu & 2.49 & Display a short view of a complex R value (list, vector, etc.) \\ 
  pipeR & 2.46 & Another pipe suite \\ 
  rethinker & 2.44 & RethinkDB client for R \\ 
   \hline
\end{tabular}
\caption{10 most polymorphic packages.}
\end{table}
